---
title: "Areas"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = F)
library(crosstalk)
library(tidyverse)
library(leaflet)
library(summarywidget)
library(DT)
library(plotly)

Areas_Test <- readRDS("Area_summary.rds") %>% 
  mutate(Habitats_P3 = ifelse(is.na(Habitats_P3), "Not paragraph 3", Habitats_P3),
         Natura_2000 = ifelse(is.na(Natura_2000), "No", Natura_2000),
         Types_markblokkort = ifelse(is.na(Types_markblokkort), "Other", Types_markblokkort))


shared_Areas <- SharedData$new(Areas_Test)
```

```{r}

filter_checkbox("Habitats_P3", "Habitats Par 3", shared_Areas, ~Habitats_P3, inline = TRUE)

filter_checkbox("Natura_2000", "Natura 2000", shared_Areas, ~Natura_2000, inline = TRUE)

filter_checkbox("Types_markblokkort", "Types markblokkort", shared_Areas, ~Types_markblokkort, inline = TRUE)

Percent <- summarywidget(shared_Areas, statistic = "sum", column = "Proportion", digits = 2, elementId = "Percent")

```

The selected areas are `r Percent` percent of Denmark

```{r}
DT::datatable(shared_Areas, extensions = 'Buttons', filter = 'top', 
            options = list(dom = 'Blfrtip',
                           scrollX='400px',
                           scrollY='200px',
                           buttons = c('copy', 'csv', 'excel',I('colvis')),
                           lengthMenu = list(c(25,50,-1),
                                             c(25,50,"All"))))  %>%
    formatRound(columns=c("Area_Sq_Mt", "Proportion"), digits=3)
```


<link rel="stylesheet" href="https://unpkg.com/leaflet@1.7.1/dist/leaflet.css" integrity="sha512-xodZBNTC5n17Xt2atTPuE1HxjVMSvLVW9ocqUKLsCC5CXdbqCmblAshOMAS6/keqq/sMZMZ19scR4PsZChSR7A==" crossorigin="" />

<script src="https://unpkg.com/leaflet@1.7.1/dist/leaflet.js" integrity="sha512-XQoYMqMTK8LvdxXYG3nZ448hOEQiglfqkJs1NOQV44cWnUrBc8PkAOcXy20w0vlaXaVUearIOBhiXZ5V3ynxwA==" crossorigin=""></script>

<script src="https://unpkg.com/georaster"></script>

<script src="https://unpkg.com/georaster-layer-for-leaflet/dist/georaster-layer-for-leaflet.min.js"></script>



```{r, warning = FALSE}
leaflet(width = 0, height = 0)
```


<div id = "map" style="width: 100%; height: 480px; position: relative;"></div>

<script>
  // Load base tiles ----

  // OpenStreetMap Tiles
  var OpenStreetMap = L.tileLayer('https://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png', {
    attribution: '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors'
  });

  // Initiate leaflet ----
  // This needs to be done before loading the Georasters
  var map = L.map('map', { 
      center: [56.25, 12], // Centered on Denmark
      zoom: 7, // Worked well
      layers: [OpenStreetMap] // Default layers
  });

  // Set URL to cog
  var random_rast_dk_url = "https://biodiversitycounciltifs.s3.eu-central-1.amazonaws.com/Paragraph3_Optimized.tif"

  //  Load georaster layer and add to map
  parseGeoraster(random_rast_dk_url).then(georaster => {
      var random_rast_dk = new GeoRasterLayer({
          georaster,
          opacity: 0.75,
          resolution: 128, // 128 seems to be a good compromise
          pixelValuesToColorFn: function(pixelValues){
              var pixelValue = pixelValues[0]; 
              if (pixelValue >= 1 && pixelValue <= 2){
                  var colour = (pixelValue == 1 ? "Purple": "DarkGreen"); 
                  return colour;
              } else {
                  return undefined;
              }   
          }
      });       
      random_rast_dk.addTo(map);
  });

</script>
