---
title: "Dataset generation for the Danish Biodiversity council"
author: "Derek Corcoran"
date: "`r format(Sys.time(), '%d/%m, %Y')`"
output:
  bookdown::github_document2:
    fig_caption: true
    toc: false
bibliography: biblio.bib
---

```{css, echo=FALSE}
.book .book-body .page-wrapper .page-inner section.normal {
  counter-reset: counter-rchunks;
}

.book .book-body .page-wrapper .page-inner section.normal div.section div.sourceCode {
  padding-right: 5em;
}

.book .book-body .page-wrapper .page-inner section.normal div.section div.sourceCode pre.demor {
  counter-increment: counter-rchunks;
  position: relative;
  overflow: visible;
}

.book .book-body .page-wrapper .page-inner section.normal div.section div.sourceCode pre.demor::after {
  content: '[' counter(counter-rchunks) ']';
  display: inline-block;
  position: absolute;
  right: -5em;
  color: rgb(48, 63, 159);
}
```

# Objective

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = T,
  warning = F,
  message = F,
  error = F
)
knitr::read_chunk('ReusableCode.R')
```


<!-- badges: start -->
<!-- badges: end -->


The goal of this repository is to generate the data sets necessary to evaluate the current and potential area that is strictly and partially protected in Denmark, note that all codes will be collapsed so that only the code that the reader needs to see can be extended.

# Packages needed

All calculations were done using the R software [@Cran2021], and all the geospatial transformations and calculations where performed using the terra package [@Hijmans2022]. There are some datasets extracted from the geodata package [@Hijmans2021]. The magrittr package was use for piping [@Milton2020], and finally the SF package was used to generate the Cloud Optimized Rasters [@Pebesma2018].

<details style=\"margin-bottom:10px;\">
<summary>Load packages</summary>
```{r LoadPackages}

```
</details>



Sea and land are treated separately

* How much is 30% and 10% according to EU
  - All natura 2.000 is in the 30%
  
* How much is 30% and 10% according to Biodiversity counsel (Biodiversitetsradet)
  - What has to be improved for natura 2.000 can be part of the 30%

## Potential areas

* NATURA 2000
  - Habitat types
  - How much is paragraph 3
  - how much area of different categories
  
  
## For each layer

* Total Area by habitat type (SqKm or meters)
* Proportion of Denmark
* Source
* Column 10% strictly protected EU
* Column 10% strictly protected Biodiversitetsradet
* Column 30% protected EU
* Column 30% protected Biodiversitetsradet
* Intersection (matrix NxN intersections?)

## For sea

Map of preasures for each polygon

## Deadlines

* **31 may:** First figures


## Test files

* Natura 2000
* BES_NATURETYPER_SHAPE
* NATUR_VILDT_RESERVATER
* Markblokke2021 Overlap with natura 2000 and paragraph 3

## How interactive should we make this

# Session info

<details style=\"margin-bottom:10px;\">
<summary>Session info</summary>
```{r }
sessioninfo::session_info()
```
</details>

# References
